More equivalences for Open pNets\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2018-11-07 2016-Weak-FH-Bisim/WeakBisim.tex

Intuition: tau is non observable if the automaton allows any tau transition from holes, and the global transition resulting from a tau in a hole is a tau transition not changing the action state.
We define $\Id(s)$ as the identity function on variables of state s.
\begin{definition}[Non-observability of silent actions]
An open automaton $A = <J,\mathcal{S},s_0,\mathcal{T}>$ \emph{cannot observe silent actions} if and only if for all $j$ in $J$ and $s$ in $\mathcal{S}$ we have
\begin{enumerate}
\item
\[ \openrule
         {
           (j\mapsto\tau),\True,\Id(s)}
         {s \OTarrow {\tau} s}
         \in \mathcal{T}
\]
and 
\item for all $\alpha$, $j$, $s$, $\Post$ and $s'$ such that
\[ \openrule
         {
           (j\mapsto\tau),\Pred,\Post}
         {s \OTarrow {\alpha} s'}
         \in \mathcal{T} \] we have \[ \alpha=\tau\land s=s'\land \Pred=\True\land\Post=\Id(s)
\]
\TODO{question : it looks like we can have 2 holes doing a tau and observe this ; I belive this is not safe (LUDO)}
\end{enumerate}
\end{definition}
The first statement of the definition state that the open automaton must allow a hole to do a silent action at any time, and must not observe it, i.e. cannot change its internal state because a hole did a $\tau$ transition. The second statement ensures that there cannot be in the open automaton other transitions that would be able to observe a silent action from a hole.

@@

\begin{lemma}[OT decomposition]\label{lem-decompose} 
	Let $\Leaves(Q)=p_l^{l\in L_Q}$; suppose\footnote{Note that the hypotheses of the 
	lemma imply that $Q$ is 
	not a \pLTS\ but a similar lemma can be proven for a pLTS $Q$}:
	\[ P[Q]_{j_0}  
		\models
		{\openrule
			{
				\beta_j^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTarrow {\alpha}
				\ostate{s_i'^{\, i\in L}}}
		}
	\]
		with  $J\cap\Holes(Q)\neq\emptyset$ or $\exists i\in L_Q.\,s_i\neq s'_i$, i.e. $Q$ takes part in the reduction.
		 Then there exist $v_Q$, $\Pred\,'$, $\Pred\,''$, 
		$\Post\,'$, $\Post\,''$ s.t.:\\[-2ex]
		%\[
		\begin{mathpar}
		P\models{\openrule
			{
				\beta_j^{j\in (J\setminus \Holes(Q)) \cup \{j_0\}}, 
				\Pred\,',  
				\Post\,'}
			{\ostate{s_i^{i\in L\setminus L_Q}} \OTarrow {\alpha}
				\ostate{s_i'^{\,i\in L\setminus L_Q}}}
		}%\]
	\vspace{-2.2ex}\\\text{and~~}
		%\[
		Q\models{\openrule
			{
				\beta_j^{j\in J\cap\Holes(Q)}, \Pred\,'',  
				\Post\,''}
			{\ostate{s_i^{i\in L_Q}} \OTarrow {\alpha_Q}
				\ostate{s_i'^{\,i\in L_Q}}}
		}%\]
		\end{mathpar}
		and  $\Pred \iff \Pred\,'
		\land \Pred\,''\land \alpha_Q=\beta_{j_0}$, $\Post=\Post\,'\uplus 
		\Post\,''$ where $\Post\,''$ is the restriction of $\Post$ over variables of 
		$\Leaves(Q)$.
\end{lemma}

@@

Using this lemma, it is straightforward to see that....

@@

Note that, because $\AlgE\cap\AlgA=\emptyset$ and $\alpha_x\!=\!\tau$ we have 
directly\NOTE{the argument is not trivial!}:
\begin{mathpar}
    \openrule
         {
           \gamma_{j x}^{j\in J_{x}}, \Pred_{OT_x},\Post_{OT_x}}
         {t \OTWeakarrow {\tau} t_x}
\end{mathpar}
\item Concerning the middle open transition, we have
 \begin{mathpar}
     \openrule
         {
           \beta_j^{J^2},\Pred\,^2,\Post^2 }
         {s^1 \OTarrow {\alpha} s^2} 
\end{mathpar}
For each of the WOTs obtained above, we have the following.
We have $(s^1,t_x|\Pred_{s^1,t_x})\in \mathcal{R}$. Then
 by Definition \ref{def-Weak-bisim} there exist weak open transitions $\symb{WOT}_{x y}^{x\in X,y\in Y} \subseteq \WT_2$:
 \begin{mathpar}
    \openrule
         {
           \gamma_{j x y}^{j\in J_{x y}}, \Pred_{OT_{x y}},\Post_{OT_{x y}}}
         {t_x \OTWeakarrow {\alpha_{x y}} t_{x y}}
\end{mathpar}
 such that  $\forall x,\{j\in J^2|\beta^2_j\neq\tau\}=J_{x}, 
(s^2,t_{x y}|\Pred_{s'^2,t_{x y}})\in \mathcal{R}$; \\
 $\Pred_{s^1,t_x} \land \Pred\,^2\\
\hspace{1cm} \implies\!\!\! \displaystyle{\bigvee_{y\in Y}}
   \left( \forall j\in J_{x y}. \vis{\beta^2_j}=\gamma_{jx y}\Rightarrow \Pred_{OT_{x y}}\land \alpha\!=\!\alpha_{x y} \land  
     \Pred_{s^2,t_{x y}}\subst{\Post^2}\subst{\Post_{OT_{x y}}}\right)$

@@

By replacing $(\Pred_{s^1,t_x}\wedge\Pred^2)$ sub-term, so the formula becomes: \\

@@

Furthermore, the rewriting of the implication will give:\\

@@

\subsection{Composition properties}
\TODO{Ludo: in the next 2 theorems I have a doubt about the biguplus of post p : not sure they should be disjoint, but else we need to do a complex composition  of posts}
\begin{lemma}[Weak OT decomposition]\label{lem-decompose} 
	Let $\Leaves(Q)=p_l^{l\in L_Q}$; suppose\footnote{Note that the hypotheses of the 
	lemma imply that $Q$ is 
	not a \pLTS\ but a similar lemma can be proven for a pLTS $Q$}:
	\[ P[Q]_{j_0}  
		\models
		{\openrule
			{
				\gamma_j^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTWeakarrow {\alpha}
				\ostate{s_i'^{\, i\in L}}}
		}
	\]
		with  $J\cap\Holes(Q)\neq\emptyset$ or $\exists i\in L_Q.\,s_i\neq s'_i$, i.e. $Q$ takes part in the reduction.
		 Then there exist $v_Q$, $\Pred\,'$,  
		$\Post\,'$,   $\gamma_{j_0}=[\beta_0..\beta_n]$, and for all $p\in[0..n]$ there exist $\Pred_p$, $\Post_p$ and a family $\gamma_{p j}^{j\in J_p}$ and for all $p\in[0..n+1]$ $s_{p i}^{\,i\in L_Q}$. s.t.:\\[-2ex]
		%\[
		\begin{mathpar}
		P\models{\openrule
			{
				\gamma_j^{j\in (J_p\setminus \Holes(Q)) \cup \{j_0\}}, 
				\Pred\,',  
				\Post\,'}
			{\ostate{s_i^{i\in L\setminus L_Q}} \OTWeakarrow {\alpha}
				\ostate{s_i'^{\,i\in L\setminus L_Q}}}
		}%\]
	\vspace{-2.2ex}\\\text{and for all $p\in[0..n]$~~}
		%\[
		Q\models{\openrule
			{
				\gamma_{p j}^{j\in J_p}, \Pred_p,  
				\Post_p}
			{\ostate{s_{p i}^{i\in L_Q}} \OTWeakarrow {\alpha_p}
				\ostate{s_{(p+1) i}^{\,i\in L_Q}}}
		}%\]
\\
		\text{and~~}  \bigcup_{\begin{array}{l}
p\in[0..n]
\end{array}} J_{p} = J\cap\Holes(Q) \qquad
 \text{and\qquad} \forall j\in J_p,\,\gamma_j= \mybigdotcup_{p=0}^n \gamma_{p j} \\
 \text{and\quad} \forall i \in L_Q.\, s_{(n+1) i} = s_i'\land s_{0 i} = s_i\\{\hspace{-3ex}}\text{and~~} \Pred \iff \Pred\,'
		\land \!\!\bigwedge_{p=0}^n(\alpha_p=\beta_p\land \Pred_p) \qquad\text{and~}
\Post=\Post\,'\uplus\!\! \biguplus_{p=0}^n
		\Post_p  
		\end{mathpar}
where $\Post_p$ only act upon variables of 
		$\Leaves(Q)$.
\end{lemma}
\TODO{proof}

>>>>

Intuition: tau is non observable if the automaton allows any tau transition from holes, and the global transition resulting from a tau in a hole is a tau transition not changing the action state.
We define $\Id(s)$ as the identity function on variables of state s.
\begin{definition}[Non-observability of silent actions]\label{def:Non-ObsTau}
An open automaton $A = <J,\mathcal{S},s_0,\mathcal{T}>$ \emph{cannot observe silent actions} if and only if for all $j$ in $J$ and $s$ in $\mathcal{S}$ we have
\begin{enumerate}
\item
\[ \openrule
         {
           (j\mapsto\tau),\True,\Id(s)}
         {s \OTarrow {\tau} s}
         \in \mathcal{T}
\]
and 
\item for all $\alpha$, $j$, $s$, $\Post$ and $s'$ such that
\[ \openrule
         {
           (j\mapsto\tau),\Pred,\Post}
         {s \OTarrow {\alpha} s'}
         \in \mathcal{T} \] we have \[ \alpha=\tau\land s=s'\land \Pred=\True\land\Post=\Id(s)
\]
\TODO{question : it looks like we can have 2 holes doing a tau and observe this ; I belive this is not safe (LUDO)}
\end{enumerate}
\end{definition}
The first statement of the definition state that the open automaton must allow a hole to do a silent action at any time, and must not observe it, i.e. cannot change its internal state because a hole did a $\tau$ transition. The second statement ensures that there cannot be in the open automaton other transitions that would be able to observe a silent action from a hole.

@@

\begin{lemma}[OT decomposition]\label{lem-decompose} 
	Let $\Leaves(Q)=p_l^{l\in L_Q}$; suppose\footnote{Note that the hypotheses of the 
	lemma imply that $Q$ is 
	not a \pLTS\ but a similar lemma can be proven for a pLTS $Q$}:
	\[ P[Q]_{j_0}  
		\models
		{\openrule
			{
				\beta_j^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTarrow {\alpha}
				\ostate{s_i'^{\, i\in L}}}
		}
	\]
		with  $J\cap\Holes(Q)\neq\emptyset$ or $\exists i\in L_Q.\,s_i\neq s'_i$, i.e. $Q$ takes part in the reduction.
		 Then there exist $\alpha_Q$, $\Pred\,'$, $\Pred\,''$, 
		$\Post\,'$, $\Post\,''$ s.t.:\\[-2ex]
		%\[
		\begin{mathpar}
		P\models{\openrule
			{
				\beta_j^{j\in (J\setminus \Holes(Q)) \cup \{j_0\}}, 
				\Pred\,',  
				\Post\,'}
			{\ostate{s_i^{i\in L\setminus L_Q}} \OTarrow {\alpha}
				\ostate{s_i'^{\,i\in L\setminus L_Q}}}
		}%\]
	\vspace{-2.2ex}\\\text{and~~}
		%\[
		Q\models{\openrule
			{
				\beta_j^{j\in J\cap\Holes(Q)}, \Pred\,'',  
				\Post\,''}
			{\ostate{s_i^{i\in L_Q}} \OTarrow {\alpha_Q}
				\ostate{s_i'^{\,i\in L_Q}}}
		}%\]
		\end{mathpar}
		and  $\Pred \iff \Pred\,'
		\land \Pred\,''\land \alpha_Q=\beta_{j_0}$, $\Post=\Post\,'\uplus 
		\Post\,''$ where $\Post\,''$ is the restriction of $\Post$ over variables of 
		$\Leaves(Q)$.
\end{lemma}

@@

\TODO{develop or delete this sentence} Using this lemma, it is straightforward to see that....

@@

Note that, because $\AlgE\cap\AlgA=\emptyset$  (actions and expressions are disjoint) and $\alpha_x\!=\!\tau$ we have 
directly ($\alpha_x$ cannot be a variable, and cannot contain expressions/variables because $\tau$ has no parameter):
\begin{mathpar}
    \openrule
         {
           \gamma_{j x}^{j\in J_{x}}, \Pred_{OT_x},\Post_{OT_x}}
         {t \OTWeakarrow {\tau} t_x}
\end{mathpar}
\item Concerning the middle open transition, we have
 \begin{mathpar}
     \openrule
         {
           \beta_j^{J^2},\Pred\,^2,\Post^2 }
         {s^1 \OTarrow {\alpha} s^2} 
\end{mathpar}
For each of the WOTs obtained above, we have the following.
We have $(s^1,t_x|\Pred_{s^1,t_x})\in \mathcal{R}$. Then
 by Definition \ref{def-Weak-bisim} there exist weak open transitions $\symb{WOT}_{x y}^{x\in X,y\in Y} \subseteq \WT_2$:
 \begin{mathpar}
    \openrule
         {
           \gamma_{j x y}^{j\in J_{x y}}, \Pred_{OT_{x y}},\Post_{OT_{x y}}}
         {t_x \OTWeakarrow {\alpha_{x y}} t_{x y}}
\end{mathpar}
 such that  $\forall x,\{j\in J^2|\beta^2_j\neq\tau\}=J_{x}, 
(s^2,t_{x y}|\Pred_{s'^2,t_{x y}})\in \mathcal{R}$; \\
 $\Pred_{s^1,t_x} \land \Pred\,^2\\
\hspace{1cm} \implies\!\!\! \displaystyle{\bigvee_{y\in Y}}
   \left( \forall j\in J_{x y}. \vis{\beta^2_j}=\gamma_{jx y}\Rightarrow \Pred_{OT_{x y}}\land \alpha\!=\!\alpha_{x y} \land  
     \Pred_{s^2,t_{x y}}\subst{\Post^2}\subst{\Post_{OT_{x y}}}\right)$

@@

By replacing $(\Pred_{s^1,t_x}\wedge\Pred^2)$ sub-term,  the formula becomes: \\

@@

Furthermore, the rewriting of the implication  gives:\\

@@

\subsection{Composition properties}
\TODO{Ludo: in the next 2 theorems I have a doubt about the biguplus of post p : not sure they should be disjoint, but else we need to do a complex composition  of posts}
\begin{lemma}[Weak OT decomposition]\label{lem-decompose} 
	Let $\Leaves(Q)=p_l^{l\in L_Q}$; suppose\footnote{Note that the hypotheses of the 
	lemma imply that $Q$ is 
	not a \pLTS\ but a similar lemma can be proven for a pLTS $Q$}:
	\[ P[Q]_{j_0}  
		\models
		{\openrule
			{
				\gamma_j^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTWeakarrow {\alpha}
				\ostate{s_i'^{\, i\in L}}}
		}
	\]
		with  $J\cap\Holes(Q)\neq\emptyset$ or $\exists i\in L_Q.\,s_i\neq s'_i$, i.e. $Q$ takes part in the reduction.
		 Then there exist $v_Q$, $\Pred\,'$,  
		$\Post\,'$,   $\gamma_{j_0}=[\beta_0..\beta_n]$, and for all $p\in[0..n]$ there exist $\Pred_p$, $\Post_p$ and a family $\gamma_{p j}^{j\in J_p}$ and for all $p\in[0..n+1]$ $s_{p i}^{\,i\in L_Q}$. s.t.:\\[-2ex]
		%\[
		\begin{mathpar}
		P\models{\openrule
			{
				\gamma_j^{j\in (J_p\setminus \Holes(Q)) \cup \{j_0\}}, 
				\Pred\,',  
				\Post\,'}
			{\ostate{s_i^{i\in L\setminus L_Q}} \OTWeakarrow {\alpha}
				\ostate{s_i'^{\,i\in L\setminus L_Q}}}
		}%\]
	\vspace{-2.2ex}\\\text{and for all $p\in[0..n]$~~}
		%\[
		Q\models{\openrule
			{
				\gamma_{p j}^{j\in J_p}, \Pred_p,  
				\Post_p}
			{\ostate{s_{p i}^{i\in L_Q}} \OTWeakarrow {\alpha_p}
				\ostate{s_{(p+1) i}^{\,i\in L_Q}}}
		}%\]
\\
		\text{and~~}  \bigcup_{\begin{array}{l}
p\in[0..n]
\end{array}} J_{p} = J\cap\Holes(Q) \qquad
 \text{and\qquad} \forall j\in J_p,\,\gamma_j= \mybigdotcup_{p=0}^n \gamma_{p j} \\
 \text{and\quad} \forall i \in L_Q.\, s_{(n+1) i} = s_i'\land s_{0 i} = s_i\\{\hspace{-3ex}}\text{and~~} \Pred \iff \Pred\,'
		\land \!\!\bigwedge_{p=0}^n(\alpha_p=\beta_p\land \Pred_p) \qquad\text{and~}
\Post=\Post\,'\uplus\!\! \biguplus_{p=0}^n
		\Post_p  
		\end{mathpar}
where $\Post_p$ only act upon variables of 
		$\Leaves(Q)$.
\end{lemma}
\LUDO{A V2 by Ludo}
\begin{lemma}[WOT decomposition]\label{lem-wot-decompose} 
	Let $\Leaves(Q)=p_l^{l\in L_Q}$; suppose\footnote{Note that the hypotheses of the 
	lemma imply that $Q$ is 
	not a \pLTS\ but a similar lemma can be proven for a pLTS $Q$}:
	\[ P[Q]_{j_0}  
		\models
		{\openrule
			{
				\gamma_j^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTWeakarrow {\alpha}
				\ostate{s_i'^{\, i\in L}}}
		}
	\]
		with  $J\cap\Holes(Q)\neq\emptyset$ or $\exists i\in L_Q.\,s_i\neq s'_i$, i.e. $Q$ takes part in the reduction\footnote{Note that necessarily $\alpha_Q\neq \tau$ as soon as silent actions are not observable Definition~\ref{def:Non-ObsTau}.}.
		 Then there exist $\alpha_Q$, $\Pred\,'$, $\Pred\,''$, 
		$\Post\,'$, $\Post\,''$ s.t.:\\[-2ex]
		%\[
		\begin{mathpar}
		P\models{\openrule
			{
				\gamma_j^{j\in (J\setminus \Holes(Q)) \cup \{j_0\}}, 
				\Pred\,',  
				\Post\,'}
			{\ostate{s_i^{i\in L\setminus L_Q}} \OTWeakarrow {\alpha}
				\ostate{s_i'^{\,i\in L\setminus L_Q}}}
		}%\]
	\vspace{-2.2ex}\\\text{and~~}
		%\[
		Q\models{\openrule
			{
				\gamma_j^{j\in J\cap\Holes(Q)}, \Pred\,'',  
				\Post\,''}
			{\ostate{s_i^{i\in L_Q}} \OTarrow {\alpha_Q}
				\ostate{s_i'^{\,i\in L_Q}}}
		}%\]
		\end{mathpar}
		and  $\Pred \iff \Pred\,'
		\land \Pred\,''\land \gamma_{j_0}=[\alpha_Q]$, $\Post=\Post\,'\uplus 
		\Post\,''$ where $\Post\,''$ is the restriction of $\Post$ over variables of 
		$\Leaves(Q)$.

\TODO{proof}
